---
title: "Homework 3 Code"
author: "Matthew Schoen"
date: May 28, 2026 
source: editor 
format: pdf
---

# Part 1: Set Up 

```{r}
#| label: set-up
#| message: false

# Reading in necessary packages 
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# Storing kelp data as specified object 
kelp <- read_csv(
  here("data", "temp-kelp.csv"))

# Storing personal data as specificed object 
step_count <- read_csv(
  here("data", "MyData_HW3_Format.csv"))
```

# Part 2: Problems 

## Problem 1 - Giant Kelp Fronds

### a. An Appropriate Test

To evaluate the strength of the relationship between temperature ($C^o$) and giant kelp frond elongation rate ($cm day^-1$), a Pearson correlation test and a Spearman rank correlation test would both be suited to use. The Pearson correlation test will evaluate whether there is a linear relationship between the two continuous variables while also assuming a normal distribution and linear relationship and the Spearman-rank correlation test will evaluate if a relationship exists between the two variables without requiring a normal distribution. As both ocean temperature and elongation rate are continuous and quantitative variables, both tests could be used depending on whether the data is normally distributed and linear. 

### b. Create a Visualization 

```{r}
#| label: data-visualization

# Creating a ggplot base layer using kelp dataset 
  ggplot(data = kelp, 
         mapping = aes(x = temp_c,
                       y = kelp_elong)) +
  
  # First layer is a scatter plot to visualize individual observations 
  geom_point(size = 2,
             alpha = 0.7,
             color = "#2E8B57",
             shape = 21) + 
  # Second layer is a line of best fit using a linear regression
  geom_smooth(method = "lm",
              color = "black",
              linewidth = 1
             ) + 
  # Using custom theme 
  theme_minimal() + 
  # Relabeling x & y axes
  labs(x = expression("Ocean Temperature (C"^o~")"),
       y = expression("Kelp Elongation Rate (cm day"^-1~")"),
       title = "The Effect of Ocean Water Temperature On 
       Kelp Elongation Rate")
```

### c. Check Your Assumptions and Run Your Test 

```{r}
#| label: assumptions-test
```


### d. Results Communication 

### e. Test Implications 

### f. Double Check Your Own Work 

## Problem 2 - Personal Data 

### a. Updating Your Visualizations

### b. Captions 

## Problem 3 - Affective Visualization 

### a. Describe An Affective Visualization 

### b. Create A Sketch 

### c. Make A Draft 

### d. Write An Artist Statement 

### e. Prep Your Materials 

## Problem 4 - Statistical Critique 

### a. Revisit and Summarize 

### b. Visual Clarity 

### c. Aesthetic Clarity 

### d. Reccomendations 



